Continuous developmental change explains discontinuities in word learning

### 2020-03-21 journal/ms_blind.Rmd
more revisions

Some researchers [@swingley2007; @yoshida2009; @pajak2016] proposed that word knowledge may not be encoded in a binary fashion, i.e., it is not the case that children either succeed or fail in encoding minimal contrast when learning the meanings. Rather, they may be encoding this knowledge in a graded fashion (see @Munakata2001 for an detailed discussion of a similar view). Thus, development does not so much involve a qualitative shift (i.e., a sudden emergence of an ability that did not exist before) as much as it consists in the continuous refinement of initially noisy knowledge.

Many different computational formalisms can represent graded knowledge. Here we use probabilistic models, a formalism that allows both easy examination of internal representations and quantification of the robustness of these representations. Word knowledge can be characterized with a probability distribution over sound instances organized in a similarity space. The probability is highest at the most typical sound instance. It decreases as the instance becomes less typical. The precision of word knowledge can be characterized by whether it tolerates slightly atypical pronunciations. This tolerance is captured formally by the variance of the probability distribution: larger variance indicates higher tolerance and lower precision, whereas smaller variance indicates lower tolerance and higher precision (for an illustration, see Figure \@ref(fig:illus) top and right panels).

This general framework --- in which the precision of word knowledge is characterized with the variance of a probability distribution --- can already provide an intuitive way of thinking about several findings. In particular, unlike the binary view, the probabilistic view allows for the possibility of word knowledge being both successful and noisy. This new understanding can provide an account for the fact that children show evidence of learning in some testing condition [e.g., @yoshida2009] but not in others [e.g., @stager1997] --- depending on the precision of the measurement.

@@

In addition to explaining the difference in behavior across the Switch and the preferential looking tasks, the probabilistic account explains difference in behavior within the same task. In particular, when the labels are quite distinct in the perceptual space ("lif" vs. "neem"), the probabilistic distributions do not overlap as much as in the case of similar-sounding words (Figure \@ref(fig:illus), left). This fact means that the learners will have less tolerance for the wrong association, leading to a successful rejection in the Switch task (as was reported by @stager1997 and subsequent studies using the same paradigm). Further, distinctiveness can be enhanced even for minimally different sounds when other cues highlight their difference [@rost2009; @rost2010; @thiessen2007; @yeung09; @dautriche2015].

In this framework, developmental change can be understood as an increase in the precision (i.e., a decrease in the variance) of the probabilistic knowledge, leading to a lower overlap between the distributions of similar-sounding words (Figure \@ref(fig:illus), right).  Importantly, a more precise representation still has a non-zero variance. Thus, learning difficulties can still be induced with challenging stimuli or in cognitively demanding situations as was demonstrated in adults studies [@white2013; @pajak2016].

@@

The probabilistic account was used before to explain patterns of learning and development at the qualitative level [e.g.,@swingley2007]. However, it is crucial to have a precise computational instantiation of this account which would help us 1) test this theoretical hypothesis more directly and 2) identify the particular parameters that are the locus of developmental change.  One previous study attempted to provide such a computational instantiation [@hofer2017]. However, this previous work was designed with the goal of reproducing the results of a specific study [@pajak2016] which focused on explaining the mismatch between speech perception and word learning in adults rather than on exploring the mechanism of development.

The present work proposes a model of word-pair learning based on the probabilistic account. We tested the ability of this model to both *explain* various findings in previous experiments in both children and adults (e.g., the fact that similar words are harder to learn than different words) and to *predict* new learning patterns that have not been tested before. In particular, we test the prediction that referent similarity (i.e., the confusability of pictures referred to by novel words) should play an identical computational role to word form similarity in predicting recognition difficulty. Although this prediction is intuitive, to our knowledge, it has never been tested.  Finally, we explore the extent to which the probabilistic account allows us to understand development in terms of as a continuous refinement in similar representations across the lifespan.

The paper is organized as follows. First, we introduce the model and we explain how it allows us to characterize behavior in a word-pair-learning paradigm. Then we explore the predictions of the model through simulating its behavior across different parameter settings. Next, we quantify the extent to which the model's predictions account for human data we collected from both preschool children and adults. Finally, we discuss the results in the light of existing accounts of word development.

@@

A similar probabilistic structure was used by @lewis2013 to model concept learning, and by @hofer2017 to model spoken word learning. However, the first study assumed that the sounds are heard unambiguously, and the second assumed the concepts are observed unambiguously. In our model, we assume that both labels and concepts are categories with possibly overlapping distributions. In other words, we allow the model to account for the --- more realistic --- possibility that an instances of a given categories (e.g.,"bin") can be mistaken for an instance of a neighboring category (e.g., "din"), albeit with a smaller probability. We assume, for simplicity, that this probability is captured by a normal distribution\footnote{\textcolor{red}{We used a normal distribution to simplify the mathematical derivations. This is obviously not aimed at providing a precise representation for the word-forms (and even less so for the visual objects), but as we shall in the results section, this simplified representation was enough to quantify the confusability of a given stimulus with its neighbors, which is the quantity we needed to predict the accuracy of word-object mapping in our task.}}:

@@

Finally, assuming that the concepts' prior probability $P(C)$ is uniformly distributed,\footnote{This is a reasonable assumption in our particular case given the similarity of the concept pairs used in each naming situation in our experiment.} we obtain the following expression, where all conditional dependencies are now well defined:

@@

From the general expression \ref{eq:general}, we derive three exact analytical solutions instantiating different learning assumptions.

@@

This assumption yields the following probability function of choosing the correct object $o_i$ given the correspondding label instance $s_i$, where $i \in \{1,2\}$ (see the Appendix for the full mathematical derivation):

@@

where $\Delta s = s_2-s_1$ is the perceptual sound distance between the pair of sounds that we refer to as $s_1$ and $s_2$. These distances were measured empirically as shown in Figure \@ref(fig:stim).

@@

The second solution is derived by making the more general assumption that both the labels and the concepts are recovered with noise from the sounds and objects. We first introduce the simplifying assumption that the label-related uncertainty $\sigma_L$ and the concept-related uncertainty $\sigma_C$ are of a similar magnitude, i.e., $\sigma_C \approx \sigma_L = \sigma$. This assumption makes the prediction that the sound similarity and the object similarity impact word learning accuracy in exactly the same way. Furthermore, it allows us to study the behavior of the model with only one free parameter, an important consideration given the small number of datapoints available from any given infant experiment.

@@

### Model 3
We finally derive the third (and most general) solution which allows label- and concept-related uncertainties to vary independently.

@@

We report data from $N=$ `r N_good_ch` children ages 4-5 years from a nursery school. An additional $N=$ `r N_all_ch-N_good_ch` children participated but were removed from analyses (using preregistered exclusion criteria) because they were not above chance on the catch trials due to the challenging nature of our procedure (see below). We also report data from $N=$ `r N_good_ad` adult participants tested on Amazon Mechanical Turk. An additional $N=$ `r N_all_ad-N_good_ad` were tested but removed from analyses (again, using preregistered exclusion criteria) because they had low scores on the catch trials or because they were familiar with the non-English sound stimuli we used in the adult experiment.

### Stimuli and similarity rating
The sound stimuli were generated using the MBROLA Speech Synthesizer [@dutoit1996]. We generated three kinds of nonsense word pairs which varied in their degree of perceptual similarity to English speakers: 1) *different* pairs: "lif"/"neem" and "zem"/"doof", 2) *intermediate* pairs: "aka"/"ama" and "ada"/"aba", and 3) *similar* non-English pairs\footnote{For the intermediate/similar pairs, we made the choice of using a VCV (Vowel-Consonant-Vowel) structure in order to emphasize the phonemic difference in the minimal pairs, especially for the non-native contrasts. However, as one reviwer rightfully pointed out to us, this structure could influence word leaning performance since it is not common in English. Nevertheless, if we observed this effect, we would have seen differences in predictability between the intermediate/similar pairs on the one hand, and different pairs (which are phonotactically common in English) on the other hand. However, this was not the case, word learning accuracy was correctly predicted based on perceptual similarity alone.}: "ada"/"a\textipa{d\super h}a" (in hindi) and "a\textipa{Q}a"/"a\textipa{\textcrh}a" (in arabic). For the simialr pairs, we used non-native sounds because this was a good way to decrease perceptual similarity for adults. In this respect, we took inspiration from the work of @pajak2016. We refer the reader to that paper for a lengthy discussion of the literature on (and implications for) L2 phoneme perception.

@@

In order to validate and quantify our similarity scales, we ran a separate survey on Amazon Mechanical Turk where we asked $N=20$ adults participants to evaluate the similarity of each sound and object pair on a 7-point scale. Data are shown in Figure \@ref(fig:stim) where we scaled responses within the range [0,1] for each stimulus group. We used these data in all models as an empirical measurement of the perceptual distance between the sound pairs and the object pairs. The use of empirical measurement allows us to eliminate $\Delta s$ and $\Delta s$ as free parameters (see @frank2012 and @xu2007 for a similar strategy).

@@

In the exposure trials, participants saw two objects associated with their corresponding sounds. We presented the first object on the left side of the tablet's screen simultaneously with the corresponding sound. The second sound-object association followed on the other side of the screen after 500ms. For both objects, visual stimuli were present for the duration of the sound clip (about 800ms). During this phase, participant were asked to pay attention and try yo lean the words. In the testing trials, participants saw both objects simultaneously and heard only one sound. They completed the trial by selecting which of the two objects corresponded to the sound. The object-sound pairings were randomized across participants, as was the order of the conditions (except for the catch condition which was always placed in the middle of the testing sequence). We also randomized the on-screen position (left vs. right) of the two pictures on each testing trial.  The online experiment can be viewed at XXX

@@

In particular, while children were not different form chance in the most difficult condition when both the sounds and objects are similar . Note, however, the differe that the difference between the these two conditions was not itself

@@

We next fit the three models obtained through expressions \ref{eq:model1}, \ref{eq:model2}, and \ref{eq:model3} to the participants' responses in each age group. The predictions of the models are shown \@ref(fig:allData). The parameter estimates (for $\sigma_L$ and $\sigma_C$) as well as models' goodness to fit (i.e., measured through $R^2$) are presented in Table \@ref(tab:models). \ Remmeber that, in our model, the fitted uncertainty parameters characterize the learners' general processing skills at a given time in development. Thus, they are understood to be the same across all stimuli within a given modality.

@@

# General Discussion

This paper explored the idea that some seemingly stage-like patterns in cognitive development can be characterized in a continuous fashion. We used as a case study the seminal work of @stager1997 showing a discrepancy between children's speech perception abilities and their word learning skills. The development of this discrepancy could be understood in terms of a discrete change in word representation. But our model demonstrates that it can also be parsimoniously described as a result of continuous developmental change in the precision of children's graded word knowledge. Our model instantiates the continuous development hypothesis [@swingley2007; @yoshida2009; @pajak2016].

More precisely, our model used the assumption that the ability to distinguish two stimuli perceptually leads to accurate mapping to two distinct objects, although this mapping can be noisy. On this account, apparent patterns of failure and success in retrieval tasks (such as ours) reflect different degrees of noise in the learned representations and not necessarily the presence/absence of these representations. With this assumption, the model was capable of accurately predicting behavior (including around-chance performance in the case of the most difficult contrasts)\footnote{That is, the case when both the sounds and objects are similar for children. Figure 6 shows that the confidence intervals include the chance value of 0.5.}, thus providing a proof of concept for the continuity hypothesis. The discontinuity hypothesis  --- which posits the absence of learning and not just retrieval difficulty --- cannot be ruled out by our model, but it is less parsimonious. We propose that the continuous hypothesis should be considered as a baseline; discontinuous development should only be posited if it deviates from this baseline.

We find in the literature two broad accounts of development in the Switch task: One that suggests *direct* development of the sound representation and one that hypothesizes *indirect* development of this representation through improvement in general cognitive resources. On the first account, the sound representation becomes more precise as learners refine the boundaries of their initially ambiguous phonetic categories and as they gain more experience with the functional role of these categories [@yoshida2009; @dietrich2007; @rost2009; @rost2010; @apfelbaum2011; @werker2005]. On the second account, the precision of sound encoding in the switch task improves as a result of the maturation of more general resources like the attentional and working memory capacity [@stager1997; @werker2004; @hofer2017]. Such improvement allows older children and adults to better encode the sound details while simultaneously matching these sounds to visual objects. Indeed, one recent meta-analysis of the switch task concluded that both changing representation precision and better memory/attention play a role in developmental changes [@tsui2019].

@@

In some sense, our model is compatible with both of these accounts. Indeed, the model aims at understanding *how* (by contrast to *what*) uncertainty can influence word learning. It characterizes and measures uncertainty in a rather global/coarse fashion, without committing to any specific source. Thus, it is compatible\footnote{Note the fact that our model is compatible with several previous accounts does not necessarily mean that these accounts are themeselves compatible with one another.} with direct accounts that posit uncertainty primarily stems from noisy phonological representations, and with the indirect account which suggests uncertainty primarily stems from a limitation in the general cognitive resources, hindering precise encoding in memory of subtle sounds under challenging conditions. Note that, since our goal was not to distinguish between these accounts, the model (at least in its current format) cannot answer questions about the development of each of these sources of uncertainty separately or about their relative contribution to the global uncertainty.

@@

While most research has focused on sound representation specifically in analyzing the process of learning similar-sounding words, this work showed that the visual representation of the referent is equally important. Indeed, Model 1 --- which assumes that any visually discriminable contrast can be encoded unambiguously as separate referents --- failed to explain the data, whereas Model 2 and 3 --- which take into account visual ambiguity --- succeeded. As a consequence of this assumption, we found that just like word learning is modulated by the phonological similarity of the form, it is also modulated by the visual similarity of the semantic referents.  This result is different from previous findings on the effect of the referents' similarity on the phonological development of the wordform [@Zhao2019; @fourtassi2019; @yeung09]. Our result is about wordform-referent mapping --- a different learning problem as it involves, in particular, the encoding of visual semantic/referent knowledge about the word (see @Zhao2019 for a detailed dsicussion). That said, these two questions can be related, albeit indirectly. For instance, while we did not model the process of phonological developmental per se, we argued in our discussion above that part of the observed difference between children and adults in terms of their word-referent mapping skills can be attributed to phonological development.

@@

Our work found, among other things, that children have higher sound/phoneme uncertainty that adults do, but prior work have shown that children have good auditory encoding [@swingley2000; @swingley2002; @swingley2003]. These two findings are not in contradicton with each other. In this work, and as we explained above, the measured sound uncertainty includes various sources. One of these sources is perceptual uncertainty, i.e., the ability to perceptually distinguish a sound from other neighboring sounds (e.g., "bin" is perceptually different from "din"), and another source is categorical uncertainty, i.e., the ability to estimate phonological boundaries, e.g., "bin" and "din" straddle a phonological boundary, that is, they signal a change in meaning. The fact that children have a higher sound uncertainty than adults do,  does not necessarily mean this higher uncertainty is due to poorer perceptual encoding, only.  In fact, it is possible that much of the differential in uncertainty (compared to adults) is due to the categorical/phonological dimension which we know undergoes a process of reorganization and refinement across development. Thus, our finding that children have higher sound-related uncertainty does not contradict previous work showing children are capable of fine-grained speech perception (see @werker2005 for a similar view).

@@

# Appendix

Here we provide the full mathematical derivation the models' predictions. More specifically, we explain how the general Equation \ref{eq:general} leads to Equations \ref{eq:model1}, \ref{eq:model2}, and \ref{eq:model3} which represent our three Models.

Each of these Models instantiates an approximation of $P(o_i|s_i)$ which is the probability that, given a label instance $s_i$, a participant would choose the correct object $o_i$, as opposed to the incorrect object $o_{j}$, where $i, j \in \{1,2\}$ and $j = 1-i$.

We strat from the general expression of this probability as defined in Equation \ref{eq:general}. This expression can be normalized in our special case (i.e., a two-alternative-forced choice task) as follows:

$$P(o_i|s_i) = \frac{\sum_{C,L} P(s_i|L)P(L|C)P(o_i|C)}{\sum_{C,L}  P(s_i|L)P(L|C)P(o_i|C) + \sum_{C,L}  P(s_i|L)P(L|C)P(o_j|C)}.$$
Using the definition of $P(L|C)$ (given in Equation \ref{eq:mapping}) and the fact that we only have two possible categories in each modality ($L_1, L_2$ and $C_1, C_2$) we obtain:

@@

In what follows, we compute the explicit expression of each conditional probability in Equation \ref{eq:special}.

$P(s_i | L_i)$ is the probability that the sound $s_i$ was sampled from its correct category $L_i$. Using the fact that we assumed in Equation \ref{eq:sound} that this probability follows a normal distribution $\mathcal{N}(\mu_C, \sigma^2_C)$, we have:

@@

We use the symplifying assumption that $\sigma_{L_i} = \sigma_L$, an uncertainty common to all labels. $\sigma_L$ can be understood as characterizing general speech processing skills (at a given time in development). We obtain the following expression:

@@

$s_i$ is supposed to be the prototyical instance of $L_i$, that is, $s_i=\mu_{L_j}$, leading to $$P(s_i | L_i) = \frac{1}{\sqrt{2\pi\sigma^2_L}}.$$

@@

$P(s_i | L_j)$ is the probability that the sound $s_i$, though a perfect instance of $L_i$, could still be a mispronunced instance of the category $L_j$. The value of this probability depends on the extent to which the distributions of the categories $L_i$ and $L_j$ overlap with each other (Figure \@ref(fig:illus)). Using the normal distribution, we obtain the following expression: $$P(s_i| L_j) = \frac{1}{\sqrt{2\pi\sigma^2_L}} e^{-\frac{(s_i-s_j)^2}{2\sigma^2_L}} = \frac{1}{\sqrt{2\pi\sigma^2_L}} e^{-\frac{\Delta s^2}{2\sigma^2_L}}.$$

@@

Finally, we substitute all these expresisions in equation \ref{eq:special} above, leading to:

>>>>

Some researchers [@swingley2007; @yoshida2009; @pajak2016] proposed that word knowledge may not be encoded in a binary fashion, i.e., it is not the case that children either succeed or fail in encoding minimal contrast when learning the meanings. Rather, they may be encoding this knowledge in a graded fashion (see @Munakata2001 for a detailed discussion of a similar view). Thus, development does not so much involve a qualitative shift (i.e., a sudden emergence of an ability that did not exist before) as much as it consists in the continuous refinement of initially noisy knowledge.

Many different computational formalisms can represent graded knowledge. Here we use probabilistic models, a formalism that allows both easy examinations of internal representations and quantification of the robustness of these representations. For example, wordform knowledge can be characterized by a probability distribution in a similarity space. The probability is highest at the most typical sound instance of the word. It decreases as the instance becomes less typical. The precision of word knowledge can be characterized by whether it tolerates slightly atypical pronunciations. This tolerance is captured formally by the variance of the probability distribution: larger variance indicates higher tolerance and lower precision, whereas smaller variance indicates lower tolerance and higher precision (for an illustration, see Figure \@ref(fig:illus) top and right panels).

This general framework --- in which the precision of word knowledge is characterized with the variance of a probability distribution --- can already provide an intuitive way of thinking about several findings. In particular, unlike the binary view, the probabilistic view allows for the possibility of word knowledge being both successful and noisy. This new understanding can provide an account for the fact that children show evidence of learning in some testing conditions [e.g., @yoshida2009] but not in others [e.g., @stager1997] --- depending on the precision of the measurement.

@@

In addition to explaining the difference in behavior across the Switch and the preferential looking tasks, the probabilistic account explains differences in behavior within the same task. In particular, when the labels are quite distinct in the perceptual space ("lif" vs. "neem"), the probabilistic distributions do not overlap as much as in the case of similar-sounding words (Figure \@ref(fig:illus), left). This fact means that the learners will have less tolerance for the wrong association, leading to a successful rejection in the Switch task (as was reported by @stager1997 and subsequent studies using the same paradigm). Further, distinctiveness can be enhanced even for minimally different sounds when other cues highlight their difference [@rost2009; @rost2010; @thiessen2007; @yeung09; @dautriche2015].

In this framework, developmental change can be understood as an increase in the precision (i.e., a decrease in the variance) of the probabilistic knowledge, leading to a lower overlap between the distributions of similar-sounding words (Figure \@ref(fig:illus), right).  Importantly, a more precise representation still has a non-zero variance. Thus, learning difficulties can still be induced with challenging stimuli or in cognitively demanding situations as was demonstrated in adult studies [@white2013; @pajak2016].

@@

The probabilistic account was used before to explain patterns of learning and development at the qualitative level [e.g.,@swingley2007]. However, it is crucial to have a precise computational instantiation of this account which would help us 1) test this theoretical hypothesis more directly and 2) identify the particular parameters that are the locus of developmental change.  One previous study attempted to provide such a computational instantiation [@hofer2017]. However, this previous work aimed more at reproducing the results of a specific study [@pajak2016] which focused on explaining the mismatch between speech perception and word learning in adults rather than on exploring the mechanism of development.

The present work proposes a model of word-pair learning and development based on the probabilistic account. We tested the ability of this model to both *explain* various findings in previous experiments in both children and adults (e.g., the fact that similar words are harder to learn than different words) and to *predict* new learning patterns that have not been tested before. In particular, we test the prediction that referent similarity (i.e., the confusability of pictures referred to by novel words) should play an identical computational role to word form similarity in predicting recognition difficulty. Although this prediction is intuitive, to our knowledge, it has never been tested.  Finally, we explore the extent to which the probabilistic account allows us to understand development in terms of as a continuous refinement in similar representations across the lifespan.

The paper is organized as follows. First, we introduce the model and we explain how it allows us to characterize behavior in a word-pair-learning paradigm. Then we explore the predictions of the model through simulating its behavior across different parameter settings. Next, we quantify the extent to which the model's predictions account for human data we collected from both preschool children and adults. Finally, we discuss the results in light of existing accounts of word development.

@@

A similar probabilistic structure was used by @lewis2013 to model concept learning, and by @hofer2017 to model spoken word learning. However, the first study assumed that the sounds are heard unambiguously, and the second assumed the concepts are observed unambiguously. In our model, we make the --- more realistic --- assumption that both labels and concepts are characterized with distributions and that the distributions of neighboring categories may overlap.  In other words, we allow the model to account for the possibility that an instance of a given category (e.g.,"bih") can be mistaken for an instance of a neighboring category (e.g., "dih"), albeit with a smaller probability. We assume, for simplicity, that this probability is captured by a normal distribution\footnote{We used a normal distribution to simplify the mathematical derivations. This is obviously not aimed at providing a precise representation for the word-forms (and even less so for the visual objects), but as we shall in the results section, this simplified representation was enough to quantify the confusability of a given stimulus with its neighbors, which is the quantity we needed to predict the accuracy of word-object mapping in our task.}:

@@

Finally, assuming that the concepts' prior probability $P(C)$ is uniformly distributed,\footnote{This is a reasonable assumption in our particular case given the similarity of the concept pairs used in each naming situation in the experiment.} we obtain the following expression, where all conditional dependencies are now well defined:

@@

From the general expression \ref{eq:general}, we derive three exact analytical solutions instantiating different learning assumptions (see the Appendix for the full mathematical derivation).

@@

Using this assumption, we can simplify the general Equation \ref{eq:general} of the probability $P(o_i | s_i)$ of choosing the correct object $o_i$ given the correspondding label instance $s_i$:

@@

where $\Delta s = s_2-s_1$ is the perceptual sound distance between the pair of sounds that we refer to as $s_1$ and $s_2$.

@@

The second solution/model is derived from Equation \ref{eq:general} by making the more general assumption that both the labels and the concepts are recovered with uncertainty from the sounds and objects, but we introduce the simplifying assumption that the label-related uncertainty $\sigma_L$ and the concept-related uncertainty $\sigma_C$ are of a similar magnitude, i.e., $\sigma_C \approx \sigma_L = \sigma$. This simplifying assumption makes the prediction that the sound similarity and the object similarity impact word learning accuracy in exactly the same way. Furthermore, it allows us to study the behavior of the model with only one free parameter, an important consideration given the small number of datapoints available from any given developmental experiment.

@@

### Model 3

We finally derive the third (and most general) solution which allows label- and concept-related uncertainties to vary independently.

@@

We report data from $N=$ `r N_good_ch` children ages 4-5 years from a nursery school. An additional $N=$ `r N_all_ch-N_good_ch` children participated but were removed from analyses using a preregistered exclusion criterion of not being above chance on the catch trials.\footnote{The number of children who were excluded is a bit high for this age group. This is partly due to our adoption of a 2x2 fully within-subject design, which is relatively novel for preschoolers. The length of this procedure was demanding for children, but we adopted this design based on pilot testing that indicated that precision of measurement was critical for testing our experimental predictions.} We also report data from $N=$ `r N_good_ad` adult participants tested on Amazon Mechanical Turk. An additional $N=$ `r N_all_ad-N_good_ad` were tested but removed from analyses using preregistered exclusion criteria of either having a low score on the catch trials or being familiar with the non-English sound stimuli we used in the adult experiment.

### Stimuli and similarity rating
The sound stimuli were generated using the MBROLA Speech Synthesizer [@dutoit1996]. We generated three kinds of nonsense word pairs which varied in their degree of perceptual similarity to English speakers: 1) *different* pairs: "lif"/"neem" and "zem"/"doof", 2) *intermediate* pairs: "aka"/"ama" and "ada"/"aba", and 3) *similar* non-English pairs\footnote{For the intermediate/similar pairs, we made the choice of using a VCV (Vowel-Consonant-Vowel) structure in order to emphasize the phonemic difference in the minimal pairs, especially for the non-native contrasts. However, as one reviwer rightfully pointed out to us, this structure could influence word leaning performance since it is not common in English. Nevertheless, if we observed this effect, we would have seen differences in predictability between the intermediate/similar pairs on the one hand, and different pairs (which are phonotactically common in English) on the other hand. However, this was not the case, word learning accuracy was correctly predicted based on perceptual similarity alone.}: "ada"/"a\textipa{d\super h}a" (in hindi) and "a\textipa{Q}a"/"a\textipa{\textcrh}a" (in arabic). For the similar pairs, we used non-native sounds because this was a good way to decrease perceptual similarity for adults. In this respect, we took inspiration from the work of @pajak2016. We refer the reader to that paper for a lengthy discussion of the literature on (and implications for) L2 phoneme perception.

@@

In order to validate and quantify our similarity scales, we ran a separate survey on Amazon Mechanical Turk where we asked $N=20$ adult participants to evaluate the similarity of each sound and object pair on a 7-point scale. Data are shown in Figure \@ref(fig:stim) where we scaled responses within the range [0,1] for each stimulus group. We used these data in all models as an empirical measurement of the perceptual distance between the sound pairs and the object pairs. The use of empirical measurement allows us to eliminate $\Delta s$ and $\Delta s$ as free parameters (see @frank2012 and @xu2007 for a similar strategy).

@@

In the exposure trials, participants saw two objects associated with their corresponding sounds. We presented the first object on the left side of the tablet's screen simultaneously with the corresponding sound. The second sound-object association followed on the other side of the screen after 500ms. For both objects, visual stimuli were present for the duration of the sound clip (about 800ms). During this phase, participants were asked to pay attention and try yo lean the words. In the testing trials, participants saw both objects simultaneously and heard only one sound. They completed the trial by selecting which of the two objects corresponded to the sound. The object-sound pairings were randomized across participants, as was the order of the conditions (except for the catch condition which was always placed in the middle of the testing sequence). We also randomized the on-screen position (left vs. right) of the two pictures on each testing trial.

@@

@@

We next fit the three models obtained through expressions \ref{eq:model1}, \ref{eq:model2}, and \ref{eq:model3} to the population's mean responses in each age group, using a Nonlinear Least Squares algorithm. The predictions of the models are shown \@ref(fig:allData). The parameter estimates (for $\sigma_L$ and $\sigma_C$) as well as models' goodness to fit (i.e., measured through $R^2$) are presented in Table \@ref(tab:models).

@@

# General Discussion
This paper explored the idea that some seemingly stage-like patterns in cognitive development can be characterized in a continuous fashion. We used as a case study the seminal work of @stager1997 showing a discrepancy between children's speech perception abilities and their word learning skills. The development of this discrepancy could be understood in terms of a discrete change in word representation. But our model demonstrates that it can also be parsimoniously described as a result of continuous developmental change in the precision of children's graded word knowledge. Our model instantiates the continuous development hypothesis [@swingley2007; @yoshida2009; @pajak2016].

More precisely, our model used the assumption that the ability to distinguish two sounds perceptually leads to an accurate mapping of these sounds with two distinct objects, although this mapping can be noisy. On this account, apparent patterns of failure and success in retrieval tasks (such as ours) reflect different degrees of noise in the learned representations and not necessarily the presence/absence of these representations. With this assumption, the model was capable of accurately predicting behavior (including around-chance performance in the case of the most difficult contrasts)\footnote{That is, the case when both the sounds and objects are similar for children. Figure 6 shows that the confidence intervals include the chance value of 0.5.}, thus providing a proof of concept for the continuity hypothesis. The discontinuity hypothesis  --- which posits the absence of learning and not just retrieval difficulty --- cannot be ruled out by our model, but it is less parsimonious. We propose that the continuous hypothesis should be considered as a baseline; discontinuous development should only be posited if it deviates from this baseline.

We find in the literature two broad accounts of development in the ability to learn the meanings of similar-sounding words: One that suggests *direct* development of the sound representation and one that hypothesizes *indirect* development of this representation through improvement in general cognitive resources. On the first account, the sound representation becomes more precise as learners refine the boundaries of their initially ambiguous phonetic categories and as they gain more experience with the functional role of these categories [@yoshida2009; @dietrich2007; @rost2009; @rost2010; @apfelbaum2011; @werker2005]. On the second account, the precision of sound encoding in the switch task improves as a result of the maturation of more general resources like the attentional and working memory capacity [@stager1997; @werker2004; @hofer2017]. Such improvement allows older children and adults to better encode the sound details while simultaneously matching these sounds to visual objects. Indeed, one recent meta-analysis of the switch task concluded that both changing representation precision and better memory/attention play a role in developmental changes [@tsui2019].

@@

Our model is compatible with both of these accounts. Indeed, the model aims at understanding *how* uncertainty can influence word learning. It characterizes and measures uncertainty in a rather global/coarse fashion, without committing to any specific source. Thus, it is compatible\footnote{The fact that our model is compatible with several previous accounts does not necessarily mean that these accounts are themselves compatible with one another.} with direct accounts that posit uncertainty primarily stems from noisy phonological representations, and with the indirect account which suggests uncertainty primarily stems from a limitation in the general cognitive resources, hindering precise encoding in memory of subtle sounds under challenging conditions. Note that, since our goal was not to distinguish between these accounts, the model (at least in its current format) cannot answer questions about the development of each of these sources of uncertainty separately or about their relative contribution to the global uncertainty.

@@

While most research has focused on sound representation specifically in analyzing the process of learning similar-sounding words, this work showed that the visual representation of the referent is equally important. Indeed, Model 1 --- which assumes that any visually discriminable contrast can be encoded unambiguously as separate referents --- failed to explain the data, whereas Model 2 and 3 --- which take into account visual ambiguity --- succeeded. As a consequence of this assumption, we found that just like word learning is modulated by the phonological similarity of the form, it is also modulated by the visual similarity of the semantic referents.  This result is different from previous findings on the effect of the referents' similarity on the phonological development of the wordform [@Zhao2019; @fourtassi2019; @yeung09]. Our result is about wordform-referent mapping --- a different learning problem as it involves, in particular, the encoding of visual semantic/referent knowledge about the word (see @Zhao2019 for a detailed discussion). While we did not model the process of phonological developmental per se, we argued in our discussion above that part of the observed difference between children and adults in terms of their word-referent mapping skills can be attributed to phonological development.

@@

Our work found, among other things, that children have higher sound/phoneme uncertainty than adults do, but prior work have shown that children have good auditory encoding [@swingley2000; @swingley2002; @swingley2003]. These two findings are not in contradicton with each other. In this work, and as we explained above, the measured sound uncertainty includes various sources. One of these sources is perceptual uncertainty, i.e., the ability to perceptually distinguish a sound from other neighboring sounds (e.g., "bin" is perceptually different from "din"), and another source is categorical uncertainty, i.e., the ability to estimate phonological boundaries, e.g., "bin" and "din" straddle a phonological boundary, that is, they signal a change in meaning. The fact that children have a higher sound uncertainty than adults do,  does not necessarily mean this higher uncertainty is due to poorer perceptual encoding, only.  In fact, it is possible that much of the differential in uncertainty (compared to adults) is due to the categorical/phonological dimension which we know undergoes a process of reorganization and refinement across development. Thus, our finding that children have higher sound-related uncertainty does not contradict previous work showing children are capable of fine-grained speech perception (see @werker2005 for a similar view).

@@

# Appendix
Here we provide the full mathematical derivation of the models' predictions.

 More specifically, we explain how the general Equation \ref{eq:general} leads to Equations \ref{eq:model1}, \ref{eq:model2}, and \ref{eq:model3} which represent our three models. Each of these models instantiates an approximation of $P(o_i|s_i)$ which is the probability that, given a label instance $s_i$, a participant would choose the correct object $o_i$, as opposed to the incorrect object $o_{j}$, where $i, j \in \{1,2\}$ and $j = 1-i$.

We start with the general expression of this probability as defined in Equation \ref{eq:general}. This expression can be normalized in our special case (i.e., a two-alternative-forced choice task) as follows:

$$P(o_i|s_i) = \frac{\sum_{C,L} P(s_i|L)P(L|C)P(o_i|C)}{\sum_{C,L}  P(s_i|L)P(L|C)P(o_i|C) + \sum_{C,L}  P(s_i|L)P(L|C)P(o_j|C)}.$$
Using the definition of $P(L|C)$ (given in Equation \ref{eq:mapping}) and the fact that there are two categories in each modality ($L_1, L_2$ and $C_1, C_2$) we obtain:

@@

In what follows, we compute the explicit expression of each conditional probability in Equation \ref{eq:special}. $P(s_i | L_i)$ is the probability that the sound $s_i$ was sampled from its correct category $L_i$. This probability follows a normal distribution $\mathcal{N}(\mu_C, \sigma^2_C)$ (Equation \ref{eq:sound}), thus:

@@

We use the simplifying assumption that $\sigma_{L_i} = \sigma_L$, an uncertainty common to all labels. $\sigma_L$ can be understood as characterizing general speech processing skills (at a given time in development). We obtain the following expression:

@@

$s_i$ is supposed to be the prototypical instance of $L_i$, that is, $s_i=\mu_{L_j}$, leading to $$P(s_i | L_i) = \frac{1}{\sqrt{2\pi\sigma^2_L}}.$$

@@

$P(s_i | L_j)$ is the probability that the sound $s_i$, though a perfect instance of $L_i$, could still be a mispronounced instance of the category $L_j$. The value of this probability depends on the extent to which the distributions of the categories $L_i$ and $L_j$ overlap with each other (Figure \@ref(fig:illus)). Using the normal distribution, we obtain the following expression: $$P(s_i| L_j) = \frac{1}{\sqrt{2\pi\sigma^2_L}} e^{-\frac{(s_i-s_j)^2}{2\sigma^2_L}} = \frac{1}{\sqrt{2\pi\sigma^2_L}} e^{-\frac{\Delta s^2}{2\sigma^2_L}}.$$

@@

Finally, we substitute all these expressions in equation \ref{eq:special} above, leading to:

$$P(o_i|s_i)= \frac{1 + e^{- (\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}}{1 + e^{-(\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}+ e^{-\frac{\Delta s^2}{2\sigma_L^2}} + e^{-\frac{\Delta o^2}{2\sigma_C^2}}}$$
Which is the expression for Model 3 (Equation \ref{eq:model3}). The expression for Model 2 (Equation \ref{eq:model2}) can be deduced from that of Model 3 by positing $\sigma= \sigma_L = \sigma_C$. The expression for Model 1 (Equation \ref{eq:model1}) is also derived from Model 3 by taking the limit when $\sigma_C \rightarrow 0$.
